Fractals are Cool

### 2020-12-08 paper/paper_papaja.rmd
fixed references

  Participants viewed self paced slideshows of fractal images in sequences of iterative growth and decay phases as well as a random ordering. The time spent viewing each image was measured as a Dwell Time, following the procedures of Baldwin et al. 2019. A set of images were generated for each of fourteen different fractal types iterated from 4 to 9 times. Each participant viewed a sequential slideshow and a random slideshow. Sequential slideshows presented each fractal type through phases of growth and decay from the least complex to most complex iterations and back down from most complex to least complex. Random slideshows presented the entire set of fractal images in random order. Six participants viewed Sequential Slideshow A, starting with Fractal Type 1 to Type 14, and Random Slideshow A starting with Fractal Image 1 to Image 174. The remaining 10 participants viewed Sequential Slideshow B, starting with Fractal Type 14 to Type 1, and Random Slideshow B starting with Fractal Image 174 to Image 1. This counterbalancing was included to address the findings of previous dwell time studies in which participants dwell times systematically decrease through the slideshow. In this way, the same fractal types were not always seen at the beginning or end of the slideshow. This balancing was then reversed in analysis so that each fractal type and the corresponding dwell times could be analyzed for the entire set of 16 participants together. Fractal images were generated using Wolfram Mathematica and slideshows were programmed using PsychoPy 3.

@@

  In this study a dwell time paradigm which measures looking times of participants as they advance through a self paced slideshow, was used to measure the preferences for fractal images. Slideshows were composed of a variety of different fractal types, each of which was presented in a sequential pattern of growth and decay of complexity. Fractal images are abundant in nature, and fractal fluency theory [@taylor2006reduction] has shown that humans have evolved perceptual preferences for fractal images. Previous studies have shown that both adults and children display patterns of aesthetic preference for a particular range of fractal dimension, a mathematical quantification analogous to complexity. Work in these studies have found that preferences generally fall in a range of complexity which mimics the fractal dimension found in natural scenes. Yet up to this point in time, these studies have utilized static and independent fractal images. These studies also have generally employed an alternative forced choice paradigm to quantify participant’s preferences. This study is novel in two ways. First, we utilize the dwell time paradigm, a robust and reliable task which has been used in statistical learning and event processing studies in developmental psychology [@kosie2019attentiona]. By measuring the time participants look at, or dwell on each image, patterns in preference, curiosity, or surprise can be mapped within or across individuals as they rapidly view many images sequentially. We suggest that a dwell time paradigm provides a more nuanced insight into aesthetic preference or interest in a given stimuli set. Results from this study have verified that this is a reliable measure to capture looking patterns and preference for fractal images. The second way in which this study provides new findings to the fractal fluency literature is by utilizing dynamic fractals. As discussed earlier, fractals pervade nature, and human perception has been tuned to process and prefer images that reflect fractals in nature. But often our experiences of natural scenes are not static or independent, like the stimuli which has been used in all previous studies. To better replicate natural experiences this study utilized a slideshow of fractal images which iterate through phases of growing and decaying complexity. We sought to investigate whether participants would prefer dynamic fractals as they grow or decay in complexity. The results show that there is a significant difference in the preference for dynamic fractals in the growth and decay phases, and that participants prefer to view fractals as they grow in complexity. This can be seen in the increasing dwell times during the growth phase, and decreasing dwell times during the decay phase. What is striking is that this pattern was consistent across 14 different fractal types and across a variety of number of iterations.

>>>>

  Participants viewed self paced slideshows of fractal images in sequences of iterative growth and decay phases as well as a random ordering. The time spent viewing each image was measured as a Dwell Time, following the procedures of @hard2019attention and @kosie2019attentional. A set of images were generated for each of fourteen different fractal types iterated from 4 to 9 times. Each participant viewed a sequential slideshow and a random slideshow. Sequential slideshows presented each fractal type through phases of growth and decay from the least complex to most complex iterations and back down from most complex to least complex. Random slideshows presented the entire set of fractal images in random order. Six participants viewed Sequential Slideshow A, starting with Fractal Type 1 to Type 14, and Random Slideshow A starting with Fractal Image 1 to Image 174. The remaining 10 participants viewed Sequential Slideshow B, starting with Fractal Type 14 to Type 1, and Random Slideshow B starting with Fractal Image 174 to Image 1. This counterbalancing was included to address the findings of previous dwell time studies in which participants dwell times systematically decrease through the slideshow. In this way, the same fractal types were not always seen at the beginning or end of the slideshow. This balancing was then reversed in analysis so that each fractal type and the corresponding dwell times could be analyzed for the entire set of 16 participants together. Fractal images were generated using Wolfram Mathematica and slideshows were programmed using PsychoPy 3.

@@

  In this study a dwell time paradigm which measures looking times of participants as they advance through a self paced slideshow, was used to measure the preferences for fractal images. Slideshows were composed of a variety of different fractal types, each of which was presented in a sequential pattern of growth and decay of complexity. Fractal images are abundant in nature, and fractal fluency theory [@taylor2006reduction] has shown that humans have evolved perceptual preferences for fractal images. Previous studies have shown that both adults and children display patterns of aesthetic preference for a particular range of fractal dimension, a mathematical quantification analogous to complexity. Work in these studies have found that preferences generally fall in a range of complexity which mimics the fractal dimension found in natural scenes. Yet up to this point in time, these studies have utilized static and independent fractal images. These studies also have generally employed an alternative forced choice paradigm to quantify participant’s preferences. This study is novel in two ways. First, we utilize the dwell time paradigm, a robust and reliable task which has been used in statistical learning and event processing studies in developmental psychology [@kosie2019attentional]. By measuring the time participants look at, or dwell on each image, patterns in preference, curiosity, or surprise can be mapped within or across individuals as they rapidly view many images sequentially. We suggest that a dwell time paradigm provides a more nuanced insight into aesthetic preference or interest in a given stimuli set. Results from this study have verified that this is a reliable measure to capture looking patterns and preference for fractal images. The second way in which this study provides new findings to the fractal fluency literature is by utilizing dynamic fractals. As discussed earlier, fractals pervade nature, and human perception has been tuned to process and prefer images that reflect fractals in nature. But often our experiences of natural scenes are not static or independent, like the stimuli which has been used in all previous studies. To better replicate natural experiences this study utilized a slideshow of fractal images which iterate through phases of growing and decaying complexity. We sought to investigate whether participants would prefer dynamic fractals as they grow or decay in complexity. The results show that there is a significant difference in the preference for dynamic fractals in the growth and decay phases, and that participants prefer to view fractals as they grow in complexity. This can be seen in the increasing dwell times during the growth phase, and decreasing dwell times during the decay phase. What is striking is that this pattern was consistent across 14 different fractal types and across a variety of number of iterations.

### 2020-12-08 paper/paper_papaja.tex
fixed references

Participants viewed self paced slideshows of fractal images in sequences of iterative growth and decay phases as well as a random ordering. The time spent viewing each image was measured as a Dwell Time, following the procedures of Baldwin et al.~2019. A set of images were generated for each of fourteen different fractal types iterated from 4 to 9 times. Each participant viewed a sequential slideshow and a random slideshow. Sequential slideshows presented each fractal type through phases of growth and decay from the least complex to most complex iterations and back down from most complex to least complex. Random slideshows presented the entire set of fractal images in random order. Six participants viewed Sequential Slideshow A, starting with Fractal Type 1 to Type 14, and Random Slideshow A starting with Fractal Image 1 to Image 174. The remaining 10 participants viewed Sequential Slideshow B, starting with Fractal Type 14 to Type 1, and Random Slideshow B starting with Fractal Image 174 to Image 1. This counterbalancing was included to address the findings of previous dwell time studies in which participants dwell times systematically decrease through the slideshow. In this way, the same fractal types were not always seen at the beginning or end of the slideshow. This balancing was then reversed in analysis so that each fractal type and the corresponding dwell times could be analyzed for the entire set of 16 participants together. Fractal images were generated using Wolfram Mathematica and slideshows were programmed using PsychoPy 3.

@@

In this study a dwell time paradigm which measures looking times of participants as they advance through a self paced slideshow, was used to measure the preferences for fractal images. Slideshows were composed of a variety of different fractal types, each of which was presented in a sequential pattern of growth and decay of complexity. Fractal images are abundant in nature, and fractal fluency theory (Taylor, 2006) has shown that humans have evolved perceptual preferences for fractal images. Previous studies have shown that both adults and children display patterns of aesthetic preference for a particular range of fractal dimension, a mathematical quantification analogous to complexity. Work in these studies have found that preferences generally fall in a range of complexity which mimics the fractal dimension found in natural scenes. Yet up to this point in time, these studies have utilized static and independent fractal images. These studies also have generally employed an alternative forced choice paradigm to quantify participant's preferences. This study is novel in two ways. First, we utilize the dwell time paradigm, a robust and reliable task which has been used in statistical learning and event processing studies in developmental psychology ({\textbf{???}}). By measuring the time participants look at, or dwell on each image, patterns in preference, curiosity, or surprise can be mapped within or across individuals as they rapidly view many images sequentially. We suggest that a dwell time paradigm provides a more nuanced insight into aesthetic preference or interest in a given stimuli set. Results from this study have verified that this is a reliable measure to capture looking patterns and preference for fractal images. The second way in which this study provides new findings to the fractal fluency literature is by utilizing dynamic fractals. As discussed earlier, fractals pervade nature, and human perception has been tuned to process and prefer images that reflect fractals in nature. But often our experiences of natural scenes are not static or independent, like the stimuli which has been used in all previous studies. To better replicate natural experiences this study utilized a slideshow of fractal images which iterate through phases of growing and decaying complexity. We sought to investigate whether participants would prefer dynamic fractals as they grow or decay in complexity. The results show that there is a significant difference in the preference for dynamic fractals in the growth and decay phases, and that participants prefer to view fractals as they grow in complexity. This can be seen in the increasing dwell times during the growth phase, and decreasing dwell times during the decay phase. What is striking is that this pattern was consistent across 14 different fractal types and across a variety of number of iterations.

@@

@@

>>>>

Participants viewed self paced slideshows of fractal images in sequences of iterative growth and decay phases as well as a random ordering. The time spent viewing each image was measured as a Dwell Time, following the procedures of Hard, Meyer, and Baldwin (2019) and Kosie and Baldwin (2019). A set of images were generated for each of fourteen different fractal types iterated from 4 to 9 times. Each participant viewed a sequential slideshow and a random slideshow. Sequential slideshows presented each fractal type through phases of growth and decay from the least complex to most complex iterations and back down from most complex to least complex. Random slideshows presented the entire set of fractal images in random order. Six participants viewed Sequential Slideshow A, starting with Fractal Type 1 to Type 14, and Random Slideshow A starting with Fractal Image 1 to Image 174. The remaining 10 participants viewed Sequential Slideshow B, starting with Fractal Type 14 to Type 1, and Random Slideshow B starting with Fractal Image 174 to Image 1. This counterbalancing was included to address the findings of previous dwell time studies in which participants dwell times systematically decrease through the slideshow. In this way, the same fractal types were not always seen at the beginning or end of the slideshow. This balancing was then reversed in analysis so that each fractal type and the corresponding dwell times could be analyzed for the entire set of 16 participants together. Fractal images were generated using Wolfram Mathematica and slideshows were programmed using PsychoPy 3.

@@

In this study a dwell time paradigm which measures looking times of participants as they advance through a self paced slideshow, was used to measure the preferences for fractal images. Slideshows were composed of a variety of different fractal types, each of which was presented in a sequential pattern of growth and decay of complexity. Fractal images are abundant in nature, and fractal fluency theory (Taylor, 2006) has shown that humans have evolved perceptual preferences for fractal images. Previous studies have shown that both adults and children display patterns of aesthetic preference for a particular range of fractal dimension, a mathematical quantification analogous to complexity. Work in these studies have found that preferences generally fall in a range of complexity which mimics the fractal dimension found in natural scenes. Yet up to this point in time, these studies have utilized static and independent fractal images. These studies also have generally employed an alternative forced choice paradigm to quantify participant's preferences. This study is novel in two ways. First, we utilize the dwell time paradigm, a robust and reliable task which has been used in statistical learning and event processing studies in developmental psychology (Kosie \& Baldwin, 2019). By measuring the time participants look at, or dwell on each image, patterns in preference, curiosity, or surprise can be mapped within or across individuals as they rapidly view many images sequentially. We suggest that a dwell time paradigm provides a more nuanced insight into aesthetic preference or interest in a given stimuli set. Results from this study have verified that this is a reliable measure to capture looking patterns and preference for fractal images. The second way in which this study provides new findings to the fractal fluency literature is by utilizing dynamic fractals. As discussed earlier, fractals pervade nature, and human perception has been tuned to process and prefer images that reflect fractals in nature. But often our experiences of natural scenes are not static or independent, like the stimuli which has been used in all previous studies. To better replicate natural experiences this study utilized a slideshow of fractal images which iterate through phases of growing and decaying complexity. We sought to investigate whether participants would prefer dynamic fractals as they grow or decay in complexity. The results show that there is a significant difference in the preference for dynamic fractals in the growth and decay phases, and that participants prefer to view fractals as they grow in complexity. This can be seen in the increasing dwell times during the growth phase, and decreasing dwell times during the decay phase. What is striking is that this pattern was consistent across 14 different fractal types and across a variety of number of iterations.

@@

\leavevmode\hypertarget{ref-hard2019attention}{}%
Hard, B. M., Meyer, M., \& Baldwin, D. (2019). Attention reorganizes as structure is detected in dynamic action. \emph{Memory \& Cognition}, \emph{47}(1), 17--32.

@@

\leavevmode\hypertarget{ref-kosie2019attentional}{}%
Kosie, J. E., \& Baldwin, D. (2019). Attentional profiles linked to event segmentation are robust to missing information. \emph{Cognitive Research: Principles and Implications}, \emph{4}(1), 1--18.
